---
title: 'Project 1: Step Activity'
author: "Jeff Naylor"
date: "March 3, 2019"
output: html_document
---

##Loading the Data

```{r}
library(tidyverse)
library(downloader)
library(readr)
library(lubridate)
# Download Dataset and Unzip
url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download(url, dest="~/Activity.zip", mode="wb") 
unzip("Activity.zip")

Activity<-read.csv("Activity.csv")
```

##What is mean total number of steps taken per day?
The mean number of steps taken per day is 9,354.
```{r}
Activity %>%
  group_by(date) %>%
  summarize(sum(steps,na.rm=TRUE)) -> StepsByDate

names(StepsByDate)[2]<-"steps"

hist(StepsByDate$steps,main="Total Daily Steps")

summary(StepsByDate$steps)
```
##What is the average daily activity pattern?
The interval with the most average steps is 835.
```{r}
Activity %>%
  group_by(interval) %>%
  summarize(mean(steps,na.rm=TRUE)) ->DailyPattern

names(DailyPattern)[2]<-"steps"

plot(DailyPattern,type="l",main="Daily Step Pattern")

DailyPattern %>%
  select(interval,steps) %>%
  arrange(desc(steps))
```
##Imputing missing values
Total Missing Values: 2,304
I imputed missing values, filling in the mean for the the interval.

```{r}
Activity %>%
  group_by(interval) %>%
  mutate(steps=ifelse(is.na(steps),mean(steps,na.rm=TRUE),steps)) -> ImputedActivity

ImputedActivity %>%
  group_by(date) %>%
  summarize(sum(steps)) ->TotalDailySteps

names(TotalDailySteps)[2]<-"steps"

hist(TotalDailySteps$steps,main="Total Daily Steps")

summary(TotalDailySteps$steps)
```
The mean number of steps increased from 9,354 to 10,766.  The median increased slightly as well, from 10,395 to 10,766.

##Are there differences in activity patterns between weekdays and weekends?
There are differences.  Weekdays have a higher average, earlier in the day, while weekends have higher averages after noon.
```{r}
ImputedActivity$date<-ymd(ImputedActivity$date)
ImputedActivity$Weekend<-ifelse(weekdays(ImputedActivity$date) %in% c("Saturday","Sunday"),"weekend","weekday")

ImputedActivity %>%
  group_by(Weekend,interval) %>%
  summarize(mean(steps)) -> FinalSteps

names(FinalSteps)[3]<-"steps"

ggplot(data=FinalSteps,aes(x=interval,y=steps))+geom_line(aes(color=Weekend))
```



